A Bidirected String Graph Model for Genome Assembly

### 2012-12-11 presentation/paper.tex
paper: minimum-cost circulation

The way these problems can be solved is to (1) place a lower bound and an upper
bound on the number of times each edge may be traversed, then (2) formulate and
solve a minimum-cost network flow problem that finds the minimum total traversal
count of edges in the graph in order for every vertex to be balanced (i.e. have
the same in-degree as out-degree), then (3) find a generalized Eulerian tour
through the bidirected graph, traversing each edge the exact number of times
that was calculated in (2) and also allowing for the tour be discontinuous.

@@

@@

Edges with sufficiently high A-statistic are very likely to be single-copy edges
and therefore should be traversed exactly one time.  The remaining edges may be
traversed any number of times (with the exact number being determined during the
network flow calculation), except for edges that originally were associated with
one or more inner vertices, in which case we require the edge be traversed at
least one time so that the reads corresponding to the inner vertices are used in
the reconstruction of the genome.  This is needed because in the absence of
foreign DNA, every read was presumably sampled from the genome.

@@

\subsection{Minimum-cost network flow}

Formally, a minimum-cost network flow problem on a directed graph takes as input
for each edge $e$ an upper bound $c(e) \ge 0$ on flow and a cost per unit flow
$v(e)$, and for each vertex $v$ a supply (positive) or demand (negative)
$b(v)$ for flow.  It produces as output a flow $x(e)$ on each edge $e$ such
that $\sum_e v(e)x(e)$ is minimized subject to $x(e) \in [0, c(e)]$ and
$\sum_{u \to v}x(u \to v) + b(v) = \sum_{v \to w} x(v \to w)$.\cite{Myers2005}.

We can roughly map our problem onto this definition assigning $c(e) = 0$ when
$e$ is definitely single-copy and $c(e) = \infty$ otherwise, $b(v)$ the sum
of the lower bounds on the traversal counts of all bidirected edges incident to
$v$ with an inward head minus the sum of the lower bounds on the traversal
counts of all edges incident to $v$ with an outward head, and $v(e) = 1$.  The
resulting traversal count $t(e)$ for a bidirected edge $e$ is then the lower
bound on the traversal count for that edge plus $x(e)$, the traversal count
assigned in the network flow calculation.  The traversal count $t(e)$ represents
the total number of times the edge may be traversed in a given assembly,
counting traversals in both directions.

There is, however, a serious complication with this network flow formulation
which was not addressed in \cite{Myers2005}.  We have a bidirected graph, not a
directed graph, so it cannot be trivially mapped to a directed network flow
problem.  A possible workaround for this is to solve directed flow on the
directed equivalent to the bidirected graph, described in
\Subsection{subsec:directed}.  The calculated traversal counts for a directed
edge $e$ between, for example, two vertices $f.E \to g.E$ is not guaranteed to
be the same as the calculated traversal count for the edge that corresponds to
the other direction of the implied bidirected edge, which would be $g.B \to f.B$
in this example.  However, gWe can take the average of the two counts and use this as the
traversal count for the corresponding bidirected edge, and this

>>>>

The way these problems can be solved is to (1) place a lower bound and an upper
bound on the number of times each edge may be traversed, then (2) formulate and
solve a minimum-cost circulation problem that finds the minimum total traversal
count of edges in the graph in order for every vertex to be balanced (i.e. have
the same in-degree as out-degree), then (3) find a generalized Eulerian tour
through the bidirected graph, traversing each edge the exact number of times
that was calculated in (2) and also allowing for the tour be discontinuous.

@@

\label{subsec:a_statistics}

@@

Edges with sufficiently high A-statistic are very likely to be single-copy edges
and therefore should be traversed exactly one time.  The remaining edges may be
traversed any number of times (with the exact number being determined during the
minimum-cost circulation calculation), except for edges that originally were
associated with one or more inner vertices, in which case we require the edge be
traversed at least one time so that the reads corresponding to the inner
vertices are used in the reconstruction of the genome.  This is needed because
in the absence of foreign DNA, every read was presumably sampled from the
genome.

@@

\subsection{Minimum-cost circulation}

Formally, a minimum-cost circulation problem on a directed graph takes as input
for each edge $e$ an lower bound $l(e) \ge 0$ and an upper bound $u(e) \ge 0$ on
flow and a cost per unit flow $v(e)$, and for each vertex $v$ a supply
(positive) or demand (negative) $b(v)$ for flow.  It produces as output a flow
$x(e)$ on each edge $e$ such that $\sum_e v(e)x(e)$ is minimized subject to
$x(e) \in [l(e), u(e)]$ and $\sum_{u \to v}x(u \to v) + b(v) = \sum_{v \to w}
x(v \to w)$ \cite{Myers2005}.  In the minimum-cost circulation problem, as
opposed to less general network flow problems, no nodes are treated specially as
sources or sinks; instead, flow must be balanced at every node $v$, but only
after taking into account the supply or demand $b(v)$ at that node.

The A-statistic calculation on each edge $e$, as described in
\Subsection{subsec:a_statistics}, allows us to assign a lower bound $l(e)$ and
an upper bound $u(e)$ on the traversal count for that edge.  Note that $u(e)$
may be $\infty$, as these bounds are very approximate.  We also have $b(v) = 0$
for all $v$, because the flow at every vertex must be balanced; this corresponds
to the walk entering each vertex the same number of times that it is exited,
making the graph as a whole be balanced.

%It produces as output a flow
%$x(e)$ on each edge $e$ such that $\sum_e v(e)x(e)$ is minimized subject to
%$x(e) \in [0, c(e)]$ and $\sum_{u \to v}x(u \to v) + b(v) = \sum_{v \to w} x(v
%\to w)$.\cite{Myers2005}.

%We can roughly map our problem onto this definition assigning $c(e) = 0$ when
%$e$ is definitely single-copy and $c(e) = \infty$ otherwise, $b(v)$ the sum
%of the lower bounds on the traversal counts of all bidirected edges incident to
%$v$ with an inward head minus the sum of the lower bounds on the traversal
%counts of all edges incident to $v$ with an outward head, and $v(e) = 1$.  The
%resulting traversal count $t(e)$ for a bidirected edge $e$ is then the lower
%bound on the traversal count for that edge plus $x(e)$, the traversal count
%assigned in the network flow calculation.  The traversal count $t(e)$ represents
%the total number of times the edge may be traversed in a given assembly,
%counting traversals in both directions.

There are still two serious problems with this circulation formulation that need
to be addressed.  First, we actually have a bidirected graph, not a directed
graph.\footnote{The fact that this circulation problem actually needs to be
formulated and solved on a bidirected graph is addressed in \cite{Medvedev2009},
but not in \cite{Myers2005}, even though the latter introduced the overall
fragment string graph algorithm.}

but is a.  We have a
bidirected graph, not a directed graph, so it cannot be trivially mapped to a
directed network flow problem.  A possible workaround for this is to solve
directed flow on the directed equivalent to the bidirected graph, described in
\Subsection{subsec:directed}.  The calculated traversal counts for a directed
edge $e$ between, for example, two vertices $f.E \to g.E$ is not guaranteed to
be the same as the calculated traversal count for the edge that corresponds to
the other direction of the implied bidirected edge, which would be $g.B \to f.B$
in this example.  However, we can take the average of the two counts and use
this as the traversal count for the corresponding bidirected edge, and this
